### 2024-10-09 esop25/cpj.tex
checkpoint

It is well-known that adding or subtracting a sample from a uniform
distribution in a finite field yields a value in a uniform
distribution, meaning that samples can be used as one-time-pads with
perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
In our type system for confidentiality, we aim to approximate
distributions by tracking which program variables an expression
may depend on. But we also want to capture this fundamental mechanism
of encryption.

@@

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets. 
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. \cnote{Define $\eqspre$.}
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

>>>>

It is well-known that adding or subtracting a sample from a uniform
distribution in a finite field yields a value in a uniform
distribution, meaning that samples can be used as one-time-pads with
perfect secrecy \cite{barthe2019probabilistic,darais2019language}.
In our type system for confidentiality, we aim to approximate
distributions by tracking which program variables an expression
may depend on. But we also want to capture this fundamental mechanism
of encryption. As we will show, SMT can be used in this and
other contexts to support confidentiality properties of protocols.

@@

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and shares
its secret input encoding with the ``evaluator'' by using
it as a selection bit. That is, assuming that client 2 is the garbler and
client 1 is the evaluator, we can define:
$$
\xassign{\mx{w}{1}}{\mux{\secret{w}}{\neg\flip{w}}{\flip{w}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{w}{1}}{\OT{\secret{w}}{1}{\neg\flip{w}}{\flip{w}}}{2}}
   \toeq{\prog} \models \mx{w}{1} \eop \neg\sx{w}{2} \fplus \rx{w}{2}
\end{mathpar}
This allows us to assign the same ciphertype to $\mx{w}{1}$ as
above-- that is, $\cty{\rx{w}{2}}{\setit{\sx{w}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

@@

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{\cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets:
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ if $R_1 \cap R_2 = \varnothing$ and is otherwise
  undefined.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. 
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.
